Lattice Boltzmann study of mass transfer for complicated geometries

### 2011-12-29 Text/introduction.tex
Verified numerical analytics. Probably too slow series convergence.

\subsection{Steady mass transfer}
The problem we want to address can be formulated through the following PDE:
\beq
\begin{aligned}
&\frac{\partial C}{\partial x} U(y)=D\frac{\partial^2 C}{\partial y^2}\\
&C(0,y)=0,\, C(x,0)=C_s,\, \frac{\partial C}{\partial_y }(x,\delta)=0\\
&U(y)=U_0 \Bigl(1-\bigl(\frac{y}{\delta}\bigr)^2\Bigr)
\end{aligned}
\feq

\section{Tweaking velocity}
This section does the job of increasing velocity results. The ulimate goal is to increase velocity
to about $20-50$ times for velocity while keeping dimensionless parameters the same. The
nondimensional parameter Peclet number governs the advection-diffusion equation:
\beq
Pe=\frac{U_{bubble} N_y}{D},
\feq
where $D$ is the diffusion coefficient to prescribe. As far as we want to increase velocity, that
means one needs to increase the diffusion coefficient as well. We make a few runs with velocities
$5,10,20,40$ times larger than original velocities.

>>>>

\subsection{Steady mass transfer}
The problem we want to address can be formulated through the following PDE:
\beq
\begin{aligned}
&\frac{\partial C}{\partial x} U(y)=D\frac{\partial^2 C}{\partial y^2}\\
&C(0,y)=0,\, C(x,0)=C_s,\, \frac{\partial C}{\partial y }(x,\delta)=0\\
&U(y)=U_0 \Bigl(1-\bigl(\frac{y}{\delta}\bigr)^2\Bigr)
\end{aligned}
\feq
{\color{red} This problem is different from what I solved on computer.}

In any case here is what we want to solve as the system for :
\beq
\begin{aligned}
&\frac{\partial C}{\partial x} U(y)=D\frac{\partial^2 C}{\partial y^2}\\
&C(0,y)=C_0,\, C(x,0)=C_{wall},\, \frac{\partial C}{\partial y }(x,\delta)=0\\
&U(y)=U_0 \Bigl(\frac{y}{\delta}\Bigr)^2
\end{aligned}
\feq
The system after non-dimensionalization becomes as follows:
\beq
\begin{aligned}
&\frac{\partial C}{\partial \frac{x}{H}} \frac{U_0}{H} \Bigl(\frac{y}{H}\Bigr)^2=\frac{D}{H^2}\frac{\partial^2 C}{\partial \frac{y^2}{H^2}}\\
&C(0,y)=C_0,\, C(x,0)=C_{wall},\, \frac{\partial C}{\partial y }(x,H)=0\\
\end{aligned}
\feq
After a change of variables:
\beq
\begin{aligned}
&\zeta=\frac{x}{H} \frac{D}{U_0 H}=\frac{1}{Pe}\frac{x}{H}\\
&\xi=\frac{y}{H}
\end{aligned}
\feq
, the equation with boundary conditions becomes as follows:
\beq
\begin{aligned}
&\frac{\partial C}{\partial \zeta}\xi^2=\frac{\partial^2 C}{\partial \xi^2}\\
&C(0,\xi)=C_0\\
&C(\zeta,0)=C_{wall}\\
&\frac{\partial C}{\partial \xi}(\zeta,1)=0\\
\end{aligned}
\feq 
Still the equation cannot be solved with the separation of variables as it requires the homogenuos boundary conditions. Thus, let us introduce the new variable to be $\Theta=C-C_{wall}$. The system will be changed but not considerably:
\beq
\begin{aligned}
&\frac{\partial \Theta}{\partial \zeta}=\frac{1}{\xi^2}\frac{\partial^2 \Theta}{\partial \xi^2}\\
&\Theta(0,\xi)=C_0-C_{wall}\\
&\Theta(\zeta,0)=0,\, \partial_{\xi}\Theta(\zeta,1)=0\\
\end{aligned}
\feq 
This equation can be solved by separation of variables if the solutions is represented as the multiplication of two functions $\Theta(\zeta,\xi)=X(\zeta)Y(\xi)$. The differential equation will become as follows:
\beq
\begin{aligned}
&\frac{\mathrm{d} X(\zeta)}{\mathrm{d} \zeta}Y(\xi)=\frac{X(\zeta)}{\xi^2}\frac{\mathrm{d^2}
Y(\xi)}{\mathrm{d}\xi^2}\\
&\frac{1}{X(\zeta)}\frac{\mathrm{d} X(\zeta)}{\mathrm{d} \zeta}=\frac{1}{\xi^{2} Y(\xi)}
\frac{\mathrm{d^2} Y(\xi)}{\mathrm{d}\xi^2}
\end{aligned}
\feq
The left and right parts depend on different variables. Therefore to be equal to each other both
parts equal to constant, say $-m^4$. Thus, the separation of the variables leads to two ODE
equations:
\begin{equation*}
\begin{aligned}
&\frac{\mathrm{d}X(\zeta)}{\mathrm{d}\zeta}=-m^4 X(\zeta)\\
&\frac{\mathrm{d^2}Y(\xi)}{\mathrm{d}\xi^2}+m^4 \xi^2 Y(\xi)=0\\ 
\end{aligned}
\end{equation*}
The solution of the first equation is the exponential function:
\beq
X(\zeta)=\exp(-m^4 \zeta)\\
\feq
The solutions of the second equation are two hypergeometric function which can be expressed through
the Bessel functions \cite{abramowitz}:
\beq
\begin{aligned}
&Y_1=\sqrt{\xi}J_{\frac{1}{4}}\Bigl(\frac{m^2 \xi^2}{2}\Bigr)\\
&Y_2=\sqrt{\xi}J_{-\frac{1}{4}}\Bigl(\frac{m^2 \xi^2}{2}\Bigr)\\
&Y'_1=m^2 \xi^{3/2} J_{-\frac{3}{4}}\Bigl(\frac{m^2 \xi^2}{2}\Bigr)\\
&Y'_2=m^2 \xi^{3/2} J_{\frac{3}{4}}\Bigl(\frac{m^2 \xi^2}{2}\Bigr)\\
\end{aligned}
\feq
The solution is the summation of two functions:
\begin{equation*}
Y(x)=C_1 Y_1(x)+C_2 Y_2(x). 
\end{equation*}
One can find coefficients from the boundary conditions:
\beq
\begin{aligned}
&Y(0)=0=C_1 Y_1(0)+C_2 Y_2(0)\\
&Y_1(0)=0,\,Y_2(0)=\frac{\sqrt{2}}{\sqrt{m} \Gamma(3/4)}\, C_2=0\\
&Y'(1)=0=C_1 Y'_1(1)+C_2 Y'_2(1)=C_1 Y'_1(1)\\
&J_{-\frac{3}{4}}\Bigl(\frac{m^2}{2}\Bigr)=0
\end{aligned}
\feq
Therefore one needs to find zeros of $J_{-\frac{3}{4}}\Bigl(\frac{m^2}{2}\Bigr)$. To give some
numerical values:
$m_1=1.454997085$, $m_2=2.927133004$, $m_3=3.857578101$, $m_4=4.601777732$, $m_5=5.240824067$.
Therefore a solution in a general form is represented as:
\beq
\Theta(\zeta,\xi)=\sum_m{C_m \sqrt{\xi} J_{\frac{1}{4}}\Bigl(\frac{m^2
\xi^2}{2}\Bigr)\exp(-m^4 \zeta)}
\feq
To find unknown coefficients $C_m$ one needs to substitute the expression above to the initial
condition:
\beq
\Theta(0,\xi)=C_0-C_{wall}=\sum_m{C_m \sqrt{\xi}J_{\frac{1}{4}}\Bigl(\frac{m^2
\xi^2}{2}\Bigr)}\\
\feq
Using the Stourm-Liouville theorem one can multiply by $\xi^{5/2}
J_{\frac{1}{4}}\Bigl(\frac{m^2 \xi^2}{2}\Bigr)$ and integrate left and right parts of the equation:
\beq
\begin{aligned}
&\bigl(C_0-C_{wall}\bigr) \int_{\xi=0}^{1}{\xi^{5/2} J_{\frac{1}{4}}\Bigl(\frac{m^2
\xi^2}{2}\Bigr)\mathrm{d}\xi}=C_m\int_{\xi=0}^{1}{\xi^3 J_{\frac{1}{4}}^2\Bigl(\frac{m^2
\xi^2}{2}\Bigr)\mathrm{d}\xi}\\
&C_m = (C_0-C_{wall}) \frac{\int_{\xi=0}^{1}{\xi^{5/2} J_{\frac{1}{4}}\Bigl(\frac{m^2
\xi^2}{2}\Bigr)\mathrm{d}\xi}}{\int_{\xi=0}^{1}{\xi^3 J_{\frac{1}{4}}^2\Bigl(\frac{m^2
\xi^2}{2}\Bigr)\mathrm{d}\xi}}
\end{aligned}
\feq
Therefore, the overall solution is specified as:
\beq
\begin{aligned}
&C(x,y)=C_{wall}+\sum_{m}{C_m
\sqrt{\frac{y}{H}}J_{\frac{1}{4}}\Bigl(\frac{m^2}{2}\frac{y^2}{H^2}\Bigr)\exp\Bigl(-\frac{m^4}{Pe}
\frac { x }{H}\Bigr)}\\
&C_m = (C_0-C_{wall}) \frac{\int_{\xi=0}^{1}{\xi^{5/2} J_{\frac{1}{4}}\Bigl(\frac{m^2
\xi^2}{2}\Bigr)\mathrm{d}\xi}}{\int_{\xi=0}^{1}{\xi^3 J_{\frac{1}{4}}^2\Bigl(\frac{m^2
\xi^2}{2}\Bigr)\mathrm{d}\xi}}
\end{aligned}
\feq
\section{Tweaking velocity}

This section does the job of increasing velocity results. The ulimate goal is to increase velocity
to about $20-50$ times for velocity while keeping dimensionless parameters the same. The
nondimensional parameter Peclet number governs the advection-diffusion equation:
\beq
Pe=\frac{U_{bubble} N_y}{D},
\feq
where $D$ is the diffusion coefficient to prescribe. As far as we want to increase velocity, that
means one needs to increase the diffusion coefficient as well. We make a few runs with velocities
$5,10,20,40$ times larger than original velocities.
